---
title: "Intro to Data Visualization"
author: "Vincent Lee"
date: "3/4/2017"
output: html_document
---

> ***Random Tip of the day:*** Explore your creative side, there are many different types of plots and packages that are shared among the R community. There is also an infinite number of permutations for how you can adjust your data visualization. On your spare time, checkout http://www.r-graph-gallery.com/all-graphs/.

Data visualization is one of the most effective tools for sharing large amounts of information in simple or a easy to understand manner. Data visualization is used to tell a story or to present your findings. Making visual plots can be easy with ggplot2, but in order to make great visuals you need to understand the different data types and variables. Another friendly reminder, keep a good balance between your Ink-to-Data Ratio. In other words, do not clutter your plots with irrelevent information or a rainbow of colors. Minimalism is an art form and "simplicity is the ultimate form of sophistication" - Leonardo Da Vinci.

### Getting started

First download and import ggplot2, if you haven't done so already. (Also go to the help button above to download a data viz cheatsheet).

```{r}
##install.packages("ggplot2")
library(ggplot2)
```

Now import and explore your data (in this case, we're using a preinstalled educational dataset called "mtcars")

```{r}
data <- mtcars
```

### Making your first visual

One of the hardest things when using ggplot for the first time is understanding the syntax. So let's try to break it down step by step.

```{r, eval = F}
##Execute the ggplot function on your dataset - notice how it's empty 
ggplot(data)
```

```{r}
##Add your x and y aesthetic followed your choice of graph. Remember you can always flip your x and y variables around as well as change your graph. However, as a precaution you need to make sure you're using the right data/variable type otherwise you will execute an error.
ggplot(data, aes(x = mpg, y = wt)) + geom_point()
```

```{r}
#Comparison
plot(data$mpg, data$wt)
```

```{r}
##Now add color
ggplot(data, aes(x = mpg, y = wt, col = mpg)) + geom_point()
```

```{r}
##Now add color and size
ggplot(data, aes(x = mpg, y = wt, col = mpg, size = mpg)) + geom_point()
```

```{r}
##To avoid clutter you can create a new variable. (also get rid of size)
p <- ggplot(data, aes(x = mpg, y = wt, col = mpg)) + geom_point()
p

```

```{r}
##The graph adding size as an aesthetic was kind of misleading, let's tone it down by adding a title and a regression line
p + geom_smooth(method='lm',formula=y~x) + ggtitle("My First Plot: Miles Per Gallon vs. Weight")
```

```{r}
#Comparison
plot(data$mpg, data$wt)
```

```{r, eval = F}
##Notice the error
ggplot(data, aes(x = mpg, y = wt, col = mpg)) + geom_bar()
```

```{r}
##You should change the variable
ggplot(data, aes(x = gear)) + geom_bar(fill = "slateblue")
```

### Exploring more visualizations

```{r}
## Qplots are just an extension of ggplots

qplot(factor(gear), mpg, data=mtcars, geom=c("boxplot", "jitter"), 
   fill=gear, alpha = .5, main="MPG vs Gear",
   xlab="Gears", ylab="MPG")
```

### Conclusion

I hope this brief introduction to data visualization provides a basic foundation for you to build your own visual model. If you ever get stuck remember Google is your best friend.

### Challenges

- Try plotting different types of graphs
- Try using other datasets to create your visual plots
- Try plotting a couple graphs side-by-side (or use facet)
- Write a detailed description explaining your plots and create a story
